The Correspondence Analysis of Two Mode Networks Revisited

### 2025-03-25 main-R1.tex

This paper contributes to the stream of previous work applying CA to analyze two-mode data. Nevertheless, it departs from previous efforts to highlight aspects of CA for two-mode data analysis in SNA that have not been emphasized or treated in detail before. Particularly, I clarify the similarities and differences between the dual scores computed by CA on both modes and those obtained from the eigenvector-based approach proposed by \citet{bonacich1991simultaneous} for obtaining dual centralities in two-mode networks. These linkages are made clearer by recasting both CA and the Bonacich centralities as variations of a ``reflective'' strategy for analyzing two-mode network data via iterative dual scoring \citep{yang2022comparing, yang2020birank}, which proceeds by assigning substantively meaningful scores to the nodes in each set of a two-mode network using an iterative procedure equivalent to a spectral decomposition of a proximity matrix based on the one-mode projection.

In this vein, I show how the CA approach to two-mode network analysis---re-invented in some corners of Network Science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting, van2021correspondence}--- is part of a more general toolkit for the analysis of two-mode networks. While the ``economic complexity'' interpretation of the CA scores on the first dimension has treated the resulting scores as ordinal rank measures---in the style of a centrality index---I review recent work that shows that they are better sense as partitioning the two sets of nodes into \textit{similarity classes} based on their patterns of connectivity to the \textit{most similar} nodes in the other set \citep{kovacs2010generalized, lizardo2024two}.

Rather than ordering nodes on a unidimensional axis based on connectivity---like Bonacich's \citet{bonacich1991simultaneous} eigenvector-based approach---CA extracts latent \textit{positional} information on people and groups. Particularly, there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}. As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to \textit{similar} entities in the other mode (e.g., groups), thus following a clustering logic of ``regular equivalence.'' In this way, CA can be used as an approach to indirect blockmodeling in two-mode networks. This is similar to the way CA has been developed and used in some disciplines (e.g., computer and information science), namely, as a \textit{spectral clustering} method for detecting most similar node subsets in bipartite graphs \citep{wu2022spectral}. Overall, CA emerges as a tool central to the usual social-network-analytic task of extracting underlying structure from two-mode (and multi-mode) relational data.

@@

In one of the earliest and most comprehensive efforts introducing CA for the analysis of the composition and structure of relational data, \citet[19-21]{wasserman1990correspondence} noted that the history of CA in the social and behavioral science was of invention, collective forgetting, and subsequent re-invention and independent rediscoveries, happening both within disciplines over time and across various fields of scholarly investigation throughout the twentieth century.\footnote{As we will see in the next section, the pattern of CA reinvention and independent rediscovery continues into the twenty-first century, now involving disciplines outside the social and behavioral sciences narrowly conceived, including computer science, information science, and network science.} In the rest of the paper and starting in the next section, I show how a recent re-invention of CA for the analysis of ``economic complexity''---two-mode networks of countries/regions by the products they are most adept at producing---in the now well-established field of ``Network Science'' in the twenty-first century, allows us to revisit controversies and open questions as to the meaning of CA scores in two-mode network data, and the relationship between correspondence analysis, centrality analysis, and positional analysis (e.g., the discovery of structurally similar clusters of actors in two-mode networks). Along the way, we will be able to shed light on the link between CA scores and traditional eigenvector centrality scores for two-mode networks, as well as ascertain the exact kind of similarity that is learned by estimating CA (and traditional eigenvector centrality) scores for nodes in two-mode networks along multiple dimensions.

@@

\subsection{CA and Dual Centrality Scoring in Two-Mode Networks} \label{sec:ca}
The reader may ask what the point of going through all this reflective scoring is since \citet{bonacich1991simultaneous} already developed a dual conception of scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected. Recall that, as noted earlier in that paper, Bonacich also considered CA in passing as a possible technique to assign scores to nodes in a two-mode network but (perhaps rightly) dismissed it as producing scores that could be interpreted as a kind of ``centrality,'' noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.

While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures). In fact, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective scores discussed in the previous section \citep{mealy2019interpreting}. This makes the connection between the Bonacich and the CA scoring approaches more intimate than Bonacich thought.

@@

Where $A$ is the original affiliation matrix. Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $A$.

@@

\begin{equation}
    C^B_p = \frac{1}{\lambda}\sum_{g}a_{pg}C^B_g
    \label{eq:faust1}
\end{equation}

@@

\begin{equation}
    C^B_g = \frac{1}{\lambda}\sum_{p}a_{pg}C^B_p
    \label{eq:faust2}
\end{equation}

Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores. These equations make clear that in the Bonacich dual-scoring approach, the scores assigned to people are equal to the sum of the scores of the groups they join, and the scores assigned to groups are equivalent to the sum of the scores simultaneously assigned to their members. This approach thus nicely captures the duality property since the scores of nodes in each mode are given by aggregating their connections to nodes in the other mode.

@@

This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue, which qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.

@@

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/ca-reord.png}
            \caption{}
            \label{fig:ca-reord}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
            \caption{}
            \label{fig:bon-reord1}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
            \caption{}
            \label{fig:bon-reord2}
    \end{subfigure}
    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score, (c) row and columns re-ordered according to person and group score on the second Bonacich dimension.}
    \label{fig:ca-v-bon}
\end{figure}

\section{Comparing Bonacich and CA Dual Scoring in the Southern Women Data} \label{sec:comparing}
\subsection{Row and Column Re-ordered Affiliation Matrices} 
Figures~\ref{fig:ca-reord} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual two-mode scoring $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns blocked according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).\footnote{This, in effect, a two-mode version of constructing ``a posteriori'' (exploratory or data-driven) blockmodels, in the sense of \citet{wasserman1987stochastic}, using scores from CA to re-order the original matrix, as they did in that paper for one-mode networks.} In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure \citep[34]{wasserman1990correspondence}, and the $C^B$ re-ordered affiliation matrix having a triangular structure.

Accordingly, the $C^R$ reflective scores reveal a dual-block partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic \textit{core-periphery} partitioning \citep{borgatti2000models}, with a group of highly active women (on the top half) who selectively co-participate in highly attended events (on the left-hand side) and less active women (on the bottom half) who go to less well-attended events (on the right-hand side). In fact, as \citet[p. 206]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic \citet{breiger1974duality} projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods. In fact, the two-block core-periphery partitioning of people and groups suggested by Figure~\ref{fig:bon-reord1}---separating the first eight women starting from the top from the rest and the first five events starting from the left from the rest) is identical to that shown in~\citet[206, Table 2]{everett2013dual}.

The larger point is that the standard eigenvector and CA scores extract qualitatively different information from the two-mode network, with $C^R$ geared toward a similarity-based partitioning---with the exact kind of similarity revealed by CA yet to be determined---and $C^B$ more focused on finding a core of well-connected people and groups, separated from a less well-connected periphery.

@@

Note that this differs from the usual interpretation of the CA correspondence plot, which is typically taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups), a criterion closer to structural equivalence~\citep{desposito2014comparison}, or ``first order'' similarity \citet{kovacs2010generalized}. This interpretation implies (for instance) that two people who attend many of the same events or groups with many members in common will appear close in the plot. But this (still common) interpretation is off the mark. What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account. This type of degree-normalization can be the preferred approach in some applications since people with many memberships and groups with many members will appear to be similar to many other people and groups if we go with the raw number of other-mode objects shared.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs. 2.13 and 2.14]{fouss2016algorithms}. As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} This implies that the similarities between pairs of people and groups will be weighted by the degrees of the other-mode objects that are shared between them. Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups. In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships \citep{desposito2014comparison}.

@@

The third CA dimension, shown in the y-axis in panel (d) of Figure~\ref{fig:ca} against the first dimension separation on the x-axis, by contrast, separates \{Eleanor, Pearl, Ruth, Dorothy, Verne, Myra\} and \{E8, E9\} from the rest. We know from Figure~\ref{fig:bon-reord1} that event E8, along with E7, is one of the most popularly attended events and thus contributes little to differentiate between the main partition of people and events on the first correspondence plot in panel (c), hence its location near the plot's origin. The pattern of attendance of these six women is one characterized by affiliation with these popular events combined with sparser attendance at the events that most serve to differentiate the two primary clusters. In his original paper, \citet[164]{bonacich1991simultaneous} noted that this third dimension ``corresponded closely to the centrality analysis pattern,'' an observation likely driven by the affinity between the actors who load highly on this dimensions and two of the most central events by raw popularity, \{E8, E9\}. What we can see instead is that the third CA dimension separates actors that are most ``peripheral'' to the two main clusters according to \citet{freeman2003finding}; namely, \{Eleanor, Pearl Ruth\} on one side and \{Dorothy, Verne, Myra\} on the other, from core members of the two main blocks, who are placed in the bottom of the correspondence plot in panel (c)---low scores in the third dimension---and toward the extreme left and right ends of the plot, indicating very high absolute value scores in the first dimension.

\begin{figure}
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
        \includegraphics[width=1.0\textwidth]{Plots/bon-corr-plot.png}
            \caption{Bonacich correspondence plot.}
            \label{fig:bon-corr}
    \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel. The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis.}
    \label{fig:bon-sim}
\end{figure}

\subsection{Breiger Projections and the Bonacich Correspondence Plot} \label{sec:eigplot}
Earlier, we discussed how CA is sometimes considered a technique useful to generate a correspondence plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$. But as we have seen, this is \textit{not} what CA is designed to do. Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes \citep{desposito2014comparison}. In fact, CA is better thought of as a low-dimensional approximation of the degree-weighted projection of the affiliation matrix, as given by equations~\ref{eq:dam1} and~\ref{eq:dam2}, which represents a degree-adjusted proximity matrix between pairs of people and pairs of groups after accounting for their respective activity and sizes, respectively.

@@

This suggests that the eigenvector centralities do encode a kind of similarity between pairs of people and pairs of groups; it just happens to be the unweighted similarity scores given by their raw number of shared memberships (for people) and shared members (for groups). This also means that the dual centrality eigenvector score mixes an ordinal rank criterion with a similarity criterion since people with lots of members will end up being similar to lots of other people (and the same for large groups). This also means that higher order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode other forms of similarity between pairs of people and pairs of groups, independently of the main axis dominated by overall person activity and group size.

While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice---\citep[see][for a discussion in the one-mode network case]{iacobucci2017eigenvector}---the exact meaning of this mysterious second eigenvector remains unclear, especially in the two-mode case. We can use the same ``a posteriori blockmodeling'' \citep{wasserman1990correspondence} approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue. Figure~\ref{fig:bon-reord2} shows the affiliation matrix blocked according to the \textit{second} Bonacich dimension, and Figure~\ref{fig:bon-sim} (c) shows the ``Bonacich Correspondence Plot'' with the usual eigenvector centrality score on the x-axis plotted against the second eigenvector of the Breiger projection matrix for both persons and groups on the y-axis.

We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension reveals a \textit{dual triangular} structure, revealing two secondary core and peripheries that are not encoded in either the first Bonacich or CA dimensions. Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering of a version of the projection matrices where the proximities were designed to detect structural equivalence (e.g., the cell entry in each projection matrix was the proportion of matches between each pair of people and groups).\footnote{As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that
attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is the position of Myrna/Myra's, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension.

This analysis points to a new way to interpret the second Bonacich dimension, namely, as a way to reveal secondary core-periphery partitions net of a main core-periphery partition in the two-mode network (if any exists). Thus, in the Bonacich Correspondence Plot shown in Figure~\ref{fig:bon-sim}(c), we can see that persons and groups are ordered by their ``main'' core-periphery scores from left-to-right, while the vertical segregation reveals

@@

Which is consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}. To accomplish this, \citet[540, eq. 2 and eq. 3]{jeh2002simrank} propose that we define the similarity of each pair of people $S(p, p')$  as given by:

>>>>

@@

In one of the earliest and most comprehensive efforts introducing CA for the analysis of the composition and structure of relational data, \citet[19-21]{wasserman1990correspondence} noted that the history of CA in the social and behavioral sciences was one of invention, collective forgetting, and subsequent re-invention and independent rediscoveries, happening both within disciplines over time and across various fields of scholarly investigation at roughly contemporaneous intervals throughout the twentieth century.\footnote{As we will see in the next section, the pattern of CA reinvention and independent rediscovery continues into the twenty-first century, now involving disciplines outside the social and behavioral sciences narrowly conceived, including computer science, information science, and network science.} In the rest of the paper and starting in the next section, I show how a recent re-invention of CA for the analysis of ``economic complexity''---two-mode networks of countries/regions by the products they are most adept at producing---in the now well-established field of ``Network Science'' in the twenty-first century, allows us to revisit controversies and open questions as to the meaning of CA scores in two-mode network data, and the relationship between correspondence analysis, centrality analysis, and positional analysis (e.g., the discovery of structurally similar clusters of actors in two-mode networks). Along the way, we will be able to shed light on the link between CA scores and traditional eigenvector centrality scores for two-mode networks, as well as ascertain the exact kind of similarity that is learned by estimating CA (and traditional eigenvector centrality) scores for nodes in two-mode networks along multiple dimensions.

@@

\subsection{CA and Dual Centrality Scoring in Two-Mode Networks} \label{sec:ca}
The reader may ask what the point of going through all this reflective scoring is since \citet{bonacich1991simultaneous} already developed a dual approach to scoring nodes in two-mode networks based on a very similar idea: Define the scores of entities in one mode based on the scores of entities in the other mode to which they are connected. As we have already noted, Bonacich also considered CA in passing as a possible technique to assign scores to nodes in a two-mode network but (perhaps rightly) dismissed it as producing scores that could be interpreted as a kind of ``centrality,'' noting that it was better at capturing the clique structure of women who attended different events \citeyearpar[164]{bonacich1991simultaneous}.

While using CA to capture patterns of similarity in attendance between persons and groups makes sense (although, as we will see, the exact patterns of similarity captured by CA are not reducible to ``cliques'' as traditionally understood), CA can also be used to assign meaningful scores to nodes in a social network, which in some applications may even be interpreted as a type of centrality (or at least have a specifiable mathematical relation to scores that are usually unambiguously interpreted as centrality measures). In fact, as I will show in a bit, in the two-mode case, the first CA dimension captures the limit $(q \rightarrow \infty)$ of the reflective scores discussed in the previous section \citep{mealy2019interpreting}. This makes the connection between the Bonacich and the CA scoring approaches more intimate than Bonacich thought.

@@

Where $\mathbf{A}$ is the original affiliation matrix. Equations~\ref{eq:bon1} and~\ref{eq:bon2} have the typical form of a linear eigensystem, which means the unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda_1$, after performing a singular value decomposition (SVD) of the rectangular affiliation matrix $\mathbf{A}$.

@@

\begin{equation}
    C^B_p = \frac{1}{\lambda_1}\sum_{g}a_{pg}C^B_g
    \label{eq:faust1}
\end{equation}

@@

\begin{equation}
    C^B_g = \frac{1}{\lambda_1}\sum_{p}a_{pg}C^B_p
    \label{eq:faust2}
\end{equation}

Where $\lambda_1$ is the largest (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores. These equations make clear that in the Bonacich dual-scoring approach, the scores assigned to people are equal to the sum of the scores of the groups they join, and the scores assigned to groups are equivalent to the sum of the scores simultaneously assigned to their members. This approach thus nicely captures the duality property since the scores of nodes in each mode are given by aggregating their connections to nodes in the other mode.

@@

This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode ``Breiger'' \citeyearpar{breiger1974duality} projection matrices corresponding to the first (largest) eigenvalue, which qualifies the Bonacich scores as a ``dual projection'' method for analyzing two-mode networks in the sense of \citet{everett2013dual}.

@@

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/ca-reord1.png}
            \caption{}
            \label{fig:ca-reord1}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord1.png}
            \caption{}
            \label{fig:bon-reord1}
    \end{subfigure}
     \begin{subfigure}[b]{0.45\textwidth}
        \includegraphics[width=1.0\textwidth]{Plots/bon-reord2.png}
            \caption{}
            \label{fig:bon-reord2}
    \end{subfigure}
    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to the person and group score on the first CA dimension, (b) row and columns re-ordered according to person and group Eigenvector centrality score, (c) row and columns re-ordered according to person and group score on the second Bonacich dimension.}
    \label{fig:ca-v-bon}
\end{figure}

\section{Comparing Bonacich and CA Dual Scoring in the Southern Women Data} \label{sec:comparing}
\subsection{Row and Column Re-ordered Affiliation Matrices} 
Figures~\ref{fig:ca-reord1} and~\ref{fig:bon-reord1} illustrate the key differences between the two versions of dual two-mode scoring $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns blocked according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b).\footnote{This, in effect, a two-mode version of constructing ``a posteriori'' (exploratory or data-driven) blockmodels, in the sense of \citet{wasserman1987stochastic}, using scores from CA and Bonacich scoring to re-order the original matrix, as they did in that paper for one-mode networks.} In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure \citep[34]{wasserman1990correspondence}, and the $C^B$ re-ordered affiliation matrix having a triangular structure.

Accordingly, the $C^R$ reflective scores reveal a dual-block partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic \textit{core-periphery} partitioning \citep{borgatti2000models}, with a group of highly active women (on the top half) who selectively co-participate in highly attended events (on the left-hand side) and less active women (on the bottom half) who go to less well-attended events (on the right-hand side). In fact, as \citet[p. 206]{everett2013dual} have noted, computing the eigenvector scores for both sets of nodes in the mode network from the classic \citet{breiger1974duality} projections---as in \citet{bonacich1991simultaneous}---and using them to re-order the rows and columns of the original affiliation matrix is a way to approximate using continuous scores an ideal discrete core-periphery partitioning of the two-mode data matrix, cheaper and more efficiently compared to discrete optimization methods. In fact, the two-block core-periphery partitioning of people and groups suggested by Figure~\ref{fig:bon-reord1}---separating the first eight women starting from the top from the rest and the first five events starting from the left from the rest) is substantively identical to that shown in~\citet[206, Table 2]{everett2013dual}, who previously extracted a core-periphery partitioning of the same data.

@@

Note that this differs from the usual interpretation of the CA correspondence plot, which is typically taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups), a criterion closer to structural equivalence~\citep{desposito2014comparison}, or ``first order'' similarity \citep{kovacs2010generalized}. This interpretation implies (for instance) that two people who attend many of the same events or groups with many members in common will appear close in the plot. But this (still common) interpretation is off the mark. What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account. This type of degree-normalization can be the preferred approach in some applications since people with many memberships and groups with many members will appear to be similar to many other people and groups if we go with the raw number of other-mode objects shared.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs. 2.13 and 2.14]{fouss2016algorithms}. As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} This implies that the similarities between pairs of people and groups will be weighted by the degrees of the other-mode objects that are shared between them. Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups. In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships \citep{desposito2014comparison}.

@@

The third CA dimension, shown in the y-axis in panel (d) of Figure~\ref{fig:ca} against the first dimension separation on the x-axis, by contrast, separates \{Eleanor, Pearl, Ruth, Dorothy, Verne, Myra\} and \{E8, E9\} from the rest. We know from Figure~\ref{fig:bon-reord1} that event E8, along with E7, is one of the most popularly attended events and thus contributes little to differentiate between the main partition of people and events on the first correspondence plot in panel (c), hence its location near the plot's origin. The pattern of attendance of these six women is one characterized by affiliation with these popular events combined with sparser attendance at the events that most serve to differentiate the two primary clusters. In his original paper, \citet[164]{bonacich1991simultaneous} noted that this third dimension ``corresponded closely to the centrality analysis pattern,'' an observation likely driven by the affinity between the actors who load highly on this dimension and two of the most central events by raw popularity, \{E8, E9\}. What we can see instead is that the third CA dimension separates actors that are most ``peripheral'' to the two main clusters according to \citet{freeman2003finding}; namely, \{Eleanor, Pearl Ruth\} on one side and \{Dorothy, Verne, Myra\} on the other, from core members of the two main blocks, who are placed in the bottom of the correspondence plot in panel (c)---low scores in the third dimension---and toward the extreme left and right ends of the plot, indicating very high absolute value scores in the first dimension.

\subsection{Breiger Projections and the Bonacich Correspondence Plot} \label{sec:eigplot}
Earlier, we discussed how CA is sometimes considered a technique useful to generate a correspondence plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$. But as we have seen, this is \textit{not} what CA is designed to do. Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes \citep{desposito2014comparison}. In fact, CA is better thought of as a low-dimensional approximation of the degree-weighted projection of the affiliation matrix, as given by equations~\ref{eq:dam1} and~\ref{eq:dam2}, which represents a degree-adjusted proximity matrix---or a Markov transition matrix under a different interpretation---between pairs of people and pairs of groups after accounting for their respective activity and sizes, respectively.

@@

This suggests that the eigenvector centralities do encode a kind of similarity between pairs of people and pairs of groups; it just happens to be based on the unweighted similarity scores given by the raw number of shared memberships (for people) and shared members (for groups). In this respect, the dual centrality eigenvector score mixes an ordinal rank criterion with a similarity criterion since people with lots of memberships will end up being ``similar`` to more people on the standard projection matrix by the simple fact of having more memberships (and the same for large groups with respect to members). Persons and groups high in eigenvector centrality can be thought of as being the most similar to most others, condensing the information in the Breiger projection matrix into a single ordinal rank. Also, just like with CA and the degree-weighted projections, higher-order axes of the Bonacich centrality eigendecomposition of the Breiger projection matrices should also encode forms of similarity between pairs of people and pairs of groups, net of the main axis dominated by overall person activity and group size.

\begin{figure}[ht!]
    \captionsetup[subfigure]{font=footnotesize,labelfont=footnotesize}
    \centering
        \includegraphics[width=1.0\textwidth]{Plots/bon-corr-plot.png}
            \caption{Bonacich correspondence plot.}
            \label{fig:bon-corr}
    \caption{Bonacich dual centrality similarity plots for persons (a) and groups (b) on the left and right top panel. The bottom panel (c) shows the Bonacich eigenvector centrality correspondence plot for persons and groups with scores corresponding to the standard Bonacich eigenvector centrality on the horizontal axis and scores corresponding to the second largest eigenvector of the affiliation matrix on the vertical axis, points are colored according a k-means three-cluster solution.}
    \label{fig:bon-sim}
\end{figure}

While plotting the Bonacich centrality against higher eigenvectors is not necessarily a substantively unmotivated practice \citep[see][for discussion in the one-mode network case]{iacobucci2017eigenvector} the exact meaning of the mysterious second eigenvector remains unclear in the two-mode case. We can use the same ``a posteriori blockmodeling'' approach we used before to contrast CA and the Bonacich centrality analysis to shed light on this issue \citep{wasserman1987stochastic}. Figure~\ref{fig:bon-reord2} shows the affiliation matrix blocked according to the \textit{second} Bonacich dimension, and Figure~\ref{fig:bon-sim} (c) shows the ``Bonacich Correspondence Plot'' with the usual eigenvector centrality score on the horizontal x-axis plotted against the second eigenvector of the Breiger projection matrix for both persons and groups on the vertical y-axis.

We can see from Figure~\ref{fig:bon-reord2} that the second Bonacich dimension diplays a \textit{dual triangular} structure, revealing two secondary core and peripheries not evident in the blocking according to either the first CA or standard eigenvector scores. Interestingly, this partition is almost equivalent to that obtained in \citet[Table 5]{everett2013dual} by clustering a version of the projection matrices of the Southern Women data where the proximities were designed to detect \textit{structural equivalence}.\footnote{In that application, the cell entry in each projection matrix was the \textit{proportion of matches} between each pair of people and groups. As \citet[208]{everett2013dual} define it, ``[p]roportion matching is the number of times two profiles have the same entry (either zeros or ones) divided by the length of the profile.''} In Everett and Borgatti's \citeyearpar[208]{everett2013dual} clustering solution, ``[t]he women are then split into (a) those that attend four or fewer events and consequently attend very few of the peripheral events (Pearl, Ruth, Verne, Myrna [sic], Dorothy, Olivia, Flora), and (b) those that attend core events and the first group of peripheral events (Evelyn, Laura, Theresa, Brenda, Charlotte, Francis, Eleanor), and (c) those that attend core events and the second group of peripheral events (Helen, Katherine, Sylvia, Nora).'' Note that the only disagreement between Everett and Borgatti's partition and the three-block solution suggested by Figure~\ref{fig:bon-reord2} is Myrna/Myra's position, who is assigned to Borgatti and Everett's third group in the a posteriori blocking obtained from Bonacich's second dimension. This may be partly due to the discrepant position of ``core event'' E9 along the columns of the blocked matrix which is grouped with the second group of ``peripheral'' events \{E10, E11, E12, E13, E14\} in Figure~\ref{fig:bon-reord2}.

This analysis points to a new way to interpret the second Bonacich dimension, namely, as a way to reveal secondary core-periphery partitions net of a main core-periphery partition in the two-mode network (if any exists). Furthermore, the similarities between persons and groups defined over the two-dimensional Bonacich space, shown in Figure~\ref{fig:bon-sim} (c), can thus be used to reveal \textit{structurally equivalent} clusters of persons and events similar to those obtained by \citet[Table 5]{everett2013dual}. This is confirmed by the coloring of the persons and groups in Figure~\ref{fig:bon-reord2}) which comes from a k-means cluster analysis using the first two Bonacich dimensions to define initial cluster centroids, revealing a three-cluster solution almost equivalent to that obtained by Everett and Borgatti, with the exception of Myrna/Myra's and event E9's assingment as discussed earlier. This suggests that the Bonacich approach to similarity clustering is distinct from CA in being useful for revealing structural equivalence, thus making the Bonacich dimensions a more faithful basis for a low-rank reconstruction of the original ``input data'' than CA if that is the goal of the analysis \citep{desposito2014comparison}.

@@

This definition consistent with a GRS approach \citep[see][]{kovacs2010generalized, lizardo2024two}. To accomplish this, \citet[540, eq. 2 and eq. 3]{jeh2002simrank} propose that we define the similarity of each pair of people $S(p, p')$  as given by:
